---
title: "Exercise 1 - R Fire Weather Data Science"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Applied Fire Data Science

Playing with toy data is all well and good, but often we want to play with our own fire data and explore it.

Objectives:

* Load, clean, and plot our own fire weather data

Let's first start with loading in some of the standard packages we will use later on:

```{r packageLoading, message = FALSE, warning = FALSE}
#install.packages("riem")
#library("riem")
#library("ggplot2")

library("tidyverse")
library("lubridate")
```

## Loading Data

Let's load in data from a FEMO on the fictional Skunk Prescribed Fire. Hopefully, the results won't stink:

```{r loadData, echo=TRUE}
skunkData <- read.csv("Data/SkunkFireWxObs.csv")
```

Whenever we load in data, we want to look at it. There are a number of ways to do that.

First, we can look at what the first couple of lines of the dataframe look like:
```{r headData, echo=TRUE}
head(skunkData)
```

Interesting. What kind of data types are there?

```{r strData, echo=TRUE}
str(skunkData)
```

We see a few things here. First, the "Date" column is misname. Let's rename it.

First, let's look at the first column of the dataframe:

```{r idData, echo=TRUE}
names(skunkData)[1]
```
Great, that's definitely the offending column name. Let's fix it! How can we do that?

```{r renameData, echo=TRUE}
names(skunkData)[1] <- "Date"
```

Did it work? Let's check.

```{r renameDataCheck, echo=TRUE}
names(skunkData)
```

Awesome!

## Very Important Dates

For anyone that's fought with dates in Excel, you're in luck! Dates are easier to handle (but still tricky) in R.

Look again at the `str(skunkData)`. What kind of variable is `Date`? What kind of variable is `Time`?

What if we wanted to plot these points? We would need to combine the `Date` and `Time` columns into a new column, first. We'll need to use the `paste()` function to do that.

Let's look up how to use that:

```{r learnAboutPaste, echo = TRUE, eval = FALSE}
?paste
```

Now that we know more about it, let's make a new column that combines `Date` and `Time`:

```{r combineDateTime, echo = TRUE}
skunkData$DateTime <- paste(skunkData$Date, skunkData$Time, sep = " ")

head(skunkData)

str(skunkData)
```

Great! Let's use a basic plotting function to look at our data.

```{r basePlot, echo = TRUE, warning = TRUE, message = TRUE, error=TRUE}
plot(skunkData$DateTime, skunkData$RH_Per)
```

Unfortunately, the data aren't in a time format that R recognizes and R is throwing a vague and somewhat unhelpful error. If you look above, DateTime is a `chr` object - character! To do convert them into dates, let's use the `lubridate` package.

First, we'll need to install it and activate it:

```{r installLubridate, echo = TRUE, eval=FALSE}
install.packages("lubridate")

library("lubridate")
```


There's a lot that lubridate can do, but one of the easiest helper functions allows you to take different formats of dates and times and convert them into formats that R can read. The commands follow the flow of the date/times:

* `dmy()` = "Day / Month / Year"
* `myd()` - "Month / Year / Day"
* `ymd()` = "Year / Month / Day"
* `ymd_hms()` = "Year / Month / Day / Hours / Minutes / Seconds"
* `ymd_hm()` = "Year / Month / Day / Hours / Minutes"
etc

`lubridate` automagically parses any format that matches whatever combination your dates or date/times are in, whether there are spaces, slashes, or whatever in between.

Let's use this to parse the DateTime column, making sure to define the timezone with the command "tz":

```{r fixDateTime}
skunkData$DateTime <- mdy_hm(skunkData$DateTime, tz = "MST")

str(skunkData)
```

And now let's plot it again:

```{r basePlotreDux, echo = TRUE, warning = TRUE, message = TRUE, error=TRUE}
plot(skunkData$DateTime, skunkData$RH_Per)
```


Great! It works now!

Let's save this file so that we can use it again:

```{r save}
write.csv(x = skunkData, file = "skunkData_dateAdded.csv",row.names = FALSE)
```


# Challenge

* Plot temperature from the skunk fire data
