---
title: "Exercise 2 - Downloading Fire Weather Data"
output: rmarkdown::github_document
---

# Objectives

* Plot using basic plots with ggplot2
* Download data from riem (install packages)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("riem")

```

## Turning on the libraries we will need:

Let's load the `ggplot2` library.

```{r libraries}
library("ggplot2")
library("lubridate")
```

...and load in your Skunk Fire data that you saved before. Also, fix the dates with `lubridate`, as before.

```{r reload}
skunkData_reloaded <- read.csv("Data/skunkData_dateAdded.csv")
skunkData_reloaded$DateTime <- ymd_hms(skunkData_reloaded$DateTime,tz = "MST")
```

## Plotting Your Data

Now that we know a bit about plotting, let's plot our own data.

Let's plot RH over time using lines:

```{r lineRH}

p <- ggplot(data = skunkData_reloaded, aes(x = DateTime, y = RH_Per))
p+
  geom_line()
```

Let's change the thickness of the line:

```{r lineRH_bigLine}

p+
  geom_line(size = 3)

```

Hmm. That's a bit too big. Let's go back to the default.

Let's add points:

```{r linePointRH}

p+
  geom_line()+
  geom_point()
```

We can also change the color of points

```{r linePointRHcolor}

p+
  geom_line()+
  geom_point(color = "red")
```

or the shape and size (try googling for "r shape" for ggplot2 point shape references)...

```{r linePointRHshape}

p+
  geom_line()+
  geom_point(shape = 22, size = 3, color = "blue")
```

We can also clean it up a little:


```{r linePointRHNice}

p+
  geom_line()+
  geom_point()+
  xlab("Time")+
  ylab("RH (%)")+
  labs(title = "Skunk Fire Weather Observations",
       subtitle = "This is a subtitle",
       caption = "Source: Jane Q. Femo")+
  theme_bw()
```

### Challenge

* Plot temperature with ggplot2 against time
* Change the point color to green
* Change the point shape to a triangle

## Downloading new data

### Installing riem

Let's download some nearby weather observations and plot them alongside these data.

First, let's download the `riem` R package, a wrapper for the Iowa Environmental Mesonet. It's a package that allows you to download weather data from National Weather Service Automated Surface Observing System (ASOS) stations. Basically, these are associated with airports. 

```{r installRIEM, echo = TRUE, eval = FALSE}
install.packages("riem")
library("riem")
```

Great! Hopefully no errors.

### Downloading data

We will use the `riem_measures` function, explained in documentation [here](http://ropensci.github.io/riem/reference/riem_measures.html).

The closest ASOS station to the Skunk Fire is FNL, so let's download the corresponding date:

```{r useage, echo = TRUE, eval = TRUE}
FCO <- riem_measures("FNL", date_start = "2018-09-22", date_end = "2018-09-24")
```

Great! Let's take a look at how the data are structured and plot it:

```{r basePlot}
str(FCO)
head(FCO)

plot(FCO$relh~FCO$valid)
```

Close readers may note that this is not a dataframe but a tibble. 

Now let's overlay it with our data. How can we do that?

```{r comparePlots}

p2 <- ggplot()+
  xlab("Date / Time")+
  ylab("RH")+
  geom_point(color = "black", data = FCO, aes(x = valid, y = relh))+
  geom_point(color = "red", data = skunkData_reloaded, aes(x = DateTime, y = RH_Per))+
  labs(title = "Skunk Fire Weather Observations",
       subtitle = "This is a subtitle",
       caption = "Source: Jane Q. Femo")+
  theme_bw()

p2
```

Uh oh, that looks weird. Maybe the time zones are off? Let's check it! And FIX IT!

```{r timeFix}
tz(FCO$valid)

FCO$DateTime <- with_tz(FCO$valid, "MST")
```

Let's plot this again.


```{r comparePlotsNewTZWhoDis}

p2 <- ggplot()+
  xlab("Date / Time")+
  ylab("RH")+
  geom_point(color = "black", data = FCO, aes(x = DateTime, y = relh))+
  geom_point(color = "red", data = skunkData_reloaded, aes(x = DateTime, y = RH_Per))+
  labs(title = "Skunk Fire Weather Observations",
       subtitle = "This is a subtitle",
       caption = "Source: Jane Q. Femo")+
  theme_bw()

p2
```

Cool!

How can we make this look nicer, though?

Let's use the package `cowplot`, which is a wrapper for `ggplot2` that improves some of the plotting functions. 
`cowplot` can do all kinds of nifty things. Check out the vignettes [here](https://cran.r-project.org/web/packages/cowplot/vignettes/introduction.html).

Go ahead and install it and switch the theme to `theme_cowplot()`

```{r cowplotting}
library(cowplot)

p2 <- ggplot()+
  xlab("Date / Time")+
  ylab("RH (%)")+
  geom_point(color = "black", data = FCO, aes(x = DateTime, y = relh))+
  geom_point(color = "red", data = skunkData_reloaded, aes(x = DateTime, y = RH_Per))+
  labs(title = "Skunk Fire Weather Observations",
       subtitle = "This is a subtitle",
       caption = "Source: Jane Q. Femo")+
  theme_cowplot()

p2
```

Looking good! Let's play with dates some more and fix the scale to just show one day. We will need to use the `scale_x_datetime` scale function, which takes values for limits.

This is a bit advanced but with some good googling for "POSIXct", "ggplot2", and "scale_x_datetime", I bet you can find the answer...

Stack Overflow to the rescue!

```{r fixDate}
p2 <- ggplot()+
  xlab("Date / Time")+
  ylab("RH (%)")+
  scale_x_datetime(limits = ymd_h(c("2018-09-22 08", "2018-09-22 20"),tz = "MST"))+
  geom_point(color = "black", data = FCO, aes(x = DateTime, y = relh))+
  geom_point(color = "red", data = skunkData_reloaded, aes(x = DateTime, y = RH_Per))+
  labs(title = "Skunk Fire Weather Observations",
       caption = "Source: Jane Q. Femo")+
  theme_cowplot()

p2

```


Now we're cooking with gas. Let's save this using the `ggsave` function.

```{r savingIt}
ggsave(plot = p2, filename = "SkunkFireRH.png")
```

### Challenge

* Plot the temperature from the Skunk Fire in green and the temperature (tmpf) of from the FCO airport data in red squares (shape 15)
* Change the title of the plot to "Skunk Fire Weather: On-Site and Regional Comparison"
* Save the plot as "SkunkFireTemperature.png"

# Things To Remember

* Plotting with ggplot2
* Downloading data with `riem` and visualizing it
* Plotting two different data types with ggplot2
* Saving plots with `ggsave()`