---
title: "Exercise 03 - Basic Fire Weather Analyses"
output: rmarkdown::github_document
---

# Objectives

TODO:
 
* Download two weather data sets
* Graph distributions of data
* Complete an anova

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

## Turning on the libraries we will need:

Let's turn on `ggplot2`, `cowplot`, `riem`, and `lubridate`

```{r libraries, message=FALSE,warning=FALSE}
library("ggplot2")
library("lubridate")
library("cowplot")
library("riem")
```

## Downloading Weather Data

Let's download weather data from two weather stations, one in Cortez, CO and one in Fort Collins.

```{r downloadStations}
FCO_2018 <- riem_measures("FNL", date_start = "2018-01-01", date_end = "2018-12-31")
CEZ_2018 <- riem_measures("CEZ", date_start = "2018-01-01", date_end = "2018-12-31")
```

Let's plot the distribution of the temperatures. First, we'll need to fix the timezones (as before).

```{r fixTZ}
CEZ_2018$DateTime <- with_tz(CEZ_2018$valid, "MST")
FCO_2018$DateTime <- with_tz(FCO_2018$valid, "MST")
```

Okay, let's plot the distribution of temperatures across the year:

Let's use a new geom, geom_density, which displays a smooth histogram.

```{r density1,message=FALSE,warning=FALSE}
tempPlot <- ggplot()+
  geom_density(data = FCO_2018, fill = "green", aes(x = tmpf))+
  geom_density(data = CEZ_2018, fill = "gray50", aes(x = tmpf))+
  theme_cowplot()
tempPlot
```
That's a bit tough to read though, isn't it? Let's change the transparency using the "alpha" argument.
```{r density2, message=F}
tempPlot <- ggplot()+
  geom_density(data = FCO_2018, fill = "green", alpha = .5, aes(x = tmpf))+
  geom_density(data = CEZ_2018, fill = "gray50", alpha = .5, aes(x = tmpf))+
  theme_cowplot()
tempPlot
```

Much better.

Let's look at it as boxplots.

```{r boxplot1, message=F}
tempPlot <- ggplot()+
  coord_flip()+
  geom_boxplot(data = FCO_2018, aes(x= "Fort Collins", y = tmpf))+
  geom_boxplot(data = CEZ_2018, aes(x = "Cortez", y = tmpf))+
  theme_cowplot()
tempPlot
```

You'll see we did two new things there:
1. We flipped the axes with "coord_flip"
2. Instead of mapping the x axes to variables, we mapped them to static values "Fort Collins" and "Cortez"

### Plotting Months

Let's add a new column and use the `lubridate` command `month` to identify the months of each date.

```{r months}
CEZ_2018$Month <- month(CEZ_2018$DateTime, label = TRUE)
FCO_2018$Month <- month(FCO_2018$DateTime, label = TRUE)
```

Now let's subset to JUST the July values. We'll use the `subset` function and assign the output to two new dataframes:

```{r subset}
CEZ_2018July <- subset(CEZ_2018, Month == "Jul")
FCO_2018July <- subset(FCO_2018, Month == "Jul")
```

And plot it, including `geom_jitter`:

```{r boxplot2, message=F}
tempMonthPlot <- ggplot()+
  coord_flip()+
  geom_boxplot(data = FCO_2018July, fill = "red", aes(x = "Fort Collins", y = tmpf))+
  geom_boxplot(data = CEZ_2018July, fill = "blue", aes(x = "Cortez", y = tmpf))+
    geom_jitter(data = FCO_2018July, fill = "red", aes(x = "Fort Collins", y = tmpf))+
  geom_jitter(data = CEZ_2018July, fill = "blue", aes(x = "Cortez", y = tmpf))+
  theme_cowplot()
tempMonthPlot
```

Nice!

```{r density3, message=F}
tempMonthPlot <- ggplot()+
  geom_density(data = FCO_2018July, fill = "green", alpha = .5, aes(x = tmpf))+
  geom_density(data = CEZ_2018July, fill = "gray50", alpha = .5, aes(x = tmpf))+
  theme_cowplot()
tempMonthPlot
```

### Challenge

* Create two dataframes that have February weather for Cortez and Fort Collins.
* Plot the distributions using geom_density of February weather. Change the transparency to .75 and the colors to black and red.

## Analyses

Great! Let's plot just the temperature and relative humidity from July in Cortez:

```{r plotWx}
windWxPlot <- ggplot(data = CEZ_2018July, aes(x = tmpf, y = relh))+
  geom_point()
windWxPlot
```
Looking good. Now let's fit a linear model to these data:

```{r linearModel}
tempRH <- lm(data = CEZ_2018July, formula = relh~tmpf)
summary(tempRH)
```

We should also do an anova on the fitted model:

```{r anovaLinearModel}
anova(tempRH)
```

ggplot2 can also display some models using geom_smooth with method="lm". It will default to y~x as a format, which is what we used.

```{r plotWx_model}
windWxPlot <- ggplot(data = CEZ_2018July, aes(x = tmpf, y = relh))+
  geom_point()+
  geom_smooth(method = "lm")
windWxPlot
```

### Challenge

* Subset the Fort Collins data to just January
* Plot the relationship between temperature and windspeed (sknt)
* Fit a linear model to the relationship of temperature and relative humidity in Fort Collins in January 

